Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-28 limits/ms.tex
Changed to use only en dashes.

\begin{abstract}
We use a new non-parametric gravitational modelling tool --- \Glass\ --- to
determine what quality of data (strong lensing, stellar kinematics, and/or
stellar masses) are required to measure the circularly averaged mass profile of
a lens and its shape. \Glass\ uses an under-constrained adaptive grid of mass
pixels to model the lens, searching through thousands of models to marginalise
over model uncertainties. Our key findings are as follows: (i) for pure lens
data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;
(ii) a single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape; (iii) stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts; (iv) stellar kinematics provide a robust measure of the mass at the half light radius of the stars $r_{1/2}$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and (v) if $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta$ -- an interesting quantity in its own right. Where
information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.
\end{abstract}

@@

In this paper, we use a new program --- \Glass\ (Gravitational Lensing AnalysiS Software) --- to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of several
gravitational lens produced from mock data.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory}, we
review the key elements of lensing theory, stellar population synthesis, and
stellar dynamics we will need. In \secref{sec:mockdata}, we describe our mock data. In \secref{sec:results}, we present our results from applying \Glass\ to these mock data. Finally, in \S\ref{sec:conclusions} we present our conclusions.

@@

With the stellar mass lower bound, there is a significant improvement of the reconstruction
quality shown in \figref{main results} and \figref{main results pixel-wise} for the doubles in the
steepest mock galaxies (\mockAC{} and \mockBC). This is because these models
are dominated by stars in the inner region. By contrast, the other two galaxies
 --- where the stars contribute negligibly to the potential --- are largely unaffected.

@@

We have introduced a new gravitational lens modelling tool ---\Glass--- and used
it to test the recovery of the mass profile and shape of mock strong lensing
galaxies. Our key findings are as follows:

>>>>

\begin{abstract}
We use a new non-parametric gravitational modelling tool -- \Glass\ -- to
determine what quality of data (strong lensing, stellar kinematics, and/or
stellar masses) are required to measure the circularly averaged mass profile of
a lens and its shape. \Glass\ uses an under-constrained adaptive grid of mass
pixels to model the lens, searching through thousands of models to marginalise
over model uncertainties. Our key findings are as follows: (i) for pure lens
data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;
(ii) a single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape; (iii) stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts; (iv) stellar kinematics provide a robust measure of the mass at the half light radius of the stars $r_{1/2}$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and (v) if $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta$ -- an interesting quantity in its own right. Where
information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.
\end{abstract}

@@

In this paper, we use a new program -- \Glass\ (Gravitational Lensing AnalysiS Software) -- to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of several
gravitational lens produced from mock data.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory}, we
review the key elements of lensing theory, stellar population synthesis, and
stellar dynamics we will need. In \secref{sec:mockdata}, we describe our mock data. In \secref{sec:results}, we present our results from applying \Glass\ to these mock data. Finally, in \S\ref{sec:conclusions} we present our conclusions.

@@

With the stellar mass lower bound, there is a significant improvement of the reconstruction
quality shown in \figref{main results} and \figref{main results pixel-wise} for the doubles in the
steepest mock galaxies (\mockAC{} and \mockBC). This is because these models
are dominated by stars in the inner region. By contrast, the other two galaxies
 -- where the stars contribute negligibly to the potential -- are largely unaffected.

@@

We have introduced a new gravitational lens modelling tool -- Glass\ -- and used
it to test the recovery of the mass profile and shape of mock strong lensing
galaxies. Our key findings are as follows:
